Towards a Wide-Coverage Grammar

### 2013-06-24 slides/presentation.tex
Before mock-presentation

\usepackage{cmll}

@@

TODO: Mention discourse here.

@@

    \vspace{2 mm}
    \begin{columns}[c]
      \begin{column}{0.25\textwidth}\end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
      \end{column}
      \begin{column}{0.1\textwidth}$\to$\end{column}
      \begin{column}{0.2\textwidth}
        $$(NP \limp S) \limp N \limp N$$
        $$N \limp (NP \limp S) \limp N$$
      \end{column}
      \begin{column}{0.25\textwidth}\end{column}
    \end{columns}
  \item The IG of Frigram use different constraints in tree
    descriptions.
    \begin{itemize}
    \item The deductive syntactic model of IGs is built on a wider range
      of linear logic primitives ($\otimes$, $\with$).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\frame{}

@@

\frame{}

@@

      $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
      \textcolor{red}{V(G) \to \Lambda(\Sigma_u)}$ such that:
      \begin{itemize}
      \item $T_u = t$.
      \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
      \item For all $(v,w) \in E(G)$,
        $\textcolor{green}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
      \end{itemize}
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/generative-perspective.pdf}
    \end{column}
  \end{columns}
\end{frame}

@@

  \begin{itemize}
  \item Direct embedding of IGs into ACGs based on Perrier's theorem
    conjectured not to be practical.
  \item Created a toolkit to serve as basis for more explorations or
    implementations.
  \item Formalized graphical abstract categorial grammars.
    \begin{itemize}
    \item Helps solve multiple constraints elegantly.
    \item Facilitates discussion and inquiry into
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Make a closer examination of the relationship between extrinsic
    and pangraphical languages in G-ACGs.
  \item Determine whether ACG object languages are closed on intersection
    or not.
  \item Write grammars using the G-ACG constraint architecture. TODO:
    More specific.
  \item Try to express the multiple independent polarized features of IGs
    using G-ACGs.
  \end{itemize}
TODO: Discourse perspective.
\end{frame}

>>>>

@@

  We want to build discourse representations in a compositional manner
  by using ACGs. Therefore, we need a wide-coverage ACG grammar.

@@

    \vspace{2 mm}
    \begin{columns}[c]
      \begin{column}{0.1\textwidth}\end{column}
      \begin{column}{0.4\textwidth}
        $$(NP \limp S) \limp N \limp N$$
        $$N \limp (NP \limp S) \limp N$$
      \end{column}
      \begin{column}{0.1\textwidth}$\to$\end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
      \end{column}
      \begin{column}{0.25\textwidth}\end{column}
    \end{columns}
  \item The IG of Frigram use different constraints in tree
    descriptions.
    \begin{itemize}
    \item The deductive syntactic model of IGs is built on a wider range
      of linear logic primitives ($\otimes$, $\&$).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\newcommand{\synt}[1]{C_{\textrm{#1}}}

\begin{frame}
  \begin{block}{Abstract languages}
    Languages of (abstract) terms constrained by the constants they are
    built on and their types. Defined by a \textcolor{red}{signature}
    and a \textcolor{green}{type}.

    \begin{columns}[t]
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \synt{tatous} &: \textcolor{red}{N} \\
          \synt{Laura} &: \textcolor{red}{NP}
        \end{align*}
      \end{column}
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \synt{les} &: \textcolor{red}{N \limp NP} \\
          \synt{aime} &: \textcolor{red}{NP \limp NP \limp S}
        \end{align*}
      \end{column}
    \end{columns}

    $$\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}) :
    \textcolor{green}{S}$$
  \end{block}

  \begin{block}{Object languages}
    Languages of (object) terms which are obtained by replacing the
    constants in an existing (abstract) term using a \textcolor{blue}{lexicon}.

    \begin{columns}[t]
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{tatous}) &= \textcolor{blue}{tatous} \\
          \mathcal{L}(\synt{Laura}) &= \textcolor{blue}{Laura}
        \end{align*}
      \end{column}
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{les}) &= \textcolor{blue}{\lambda^{\circ} x.\ les + x} \\
          \mathcal{L}(\synt{aime}) &= \textcolor{blue}{\lambda^{\circ} x y.\ x + aime + y}
        \end{align*}
      \end{column}
    \end{columns}

    $$\mathcal{L}(\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}))
    =_{\beta} Laura + aime + les + tatous$$
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{ACG Diagrams}

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Two different lexicons interpret the same abstract terms.
      \item Instead of referring to the two ACGs $\mathopen{<}
        \Sigma_{Synt}, \Sigma_{String}, \mathcal{L}_{syntax}, S
        \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
        \mathcal{L}_{sem}, S \mathclose{>}$, we refer to this single
        diagram.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/double-acg.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{ACG Patterns}

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{block}{Semantic ambiguities}
        \vspace{2 mm}
        \begin{center}
          \includegraphics[height=0.7\textheight]{../diagrams/parallel-over-serial.pdf}
        \end{center}
      \end{block}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{block}{Syntactic constraint}
        \vspace{2 mm}
        \begin{center}
          \includegraphics[height=0.7\textheight]{../diagrams/serial-over-parallel.pdf}
        \end{center}
      \end{block}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Composing ACG Patterns I}

  In order to stay in the scope of ACGs, our diagrams must be shaped as
  arborescences (at most one parent per node).

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Multiple Constraints}

  Implementing multiple constraints in a single signature is
  \textbf{difficult}.

  Agreement:
    \begin{align*}
      A_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}SG} \limp S \\
      A_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}PL} \limp S
    \end{align*}

  Extraction:
    \begin{align*}
      E_{aime_1} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
      E_{aime_2} &: NP\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      E_{aime_3} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
    \end{align*}

  Combined:
  \begin{small}
  \begin{align*}
      C_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
      C_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      C_{aime_3} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      C_{aime_4} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
      C_{aime_5} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      C_{aime_6} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
  \end{align*}

  \end{small}
\end{frame}

\begin{frame}
  \frametitle{Composing ACG Patterns II}

  We will need to be able to specify constraints independently if we are
  to write a manageably simple wide-coverage grammar.

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Composing ACG Patterns III}

  Therefore, we would like to able to do something like this (i.e., not
  an arborescence).

  \begin{columns}[c]
    \begin{column}{0.6\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/final.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.4\textwidth}
      How do we assign languages to nodes of the graph? \\
      $\implies$ Graphical Abstract Categorial Grammars (G-ACGs)

      \vspace{5 mm}

      Formally, we have:
        \begin{itemize}
        \item A DAG.
        \item Nodes labelled with:
          \begin{itemize}
          \item Signatures.
          \item Distinguished types.
          \end{itemize}
        \item Edges labelled with:
          \begin{itemize}
          \item Lexicons.
          \end{itemize}
        \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

@@

      $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
      \textcolor{red}{V(G) \to \Lambda}$ such that:
      \begin{itemize}
      \item $T_u = t$.
      \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
      \item For all $(v,w) \in E(G)$,
        $\textcolor{green}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
      \end{itemize}
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/generative-perspective.pdf}
    \end{column}
  \end{columns}
\end{frame}

@@

  \begin{itemize}
  \item Direct embedding of IGs into ACGs based on Perrier's theorem
    conjectured not to be practical.
  \item Created a toolkit to serve as basis for more explorations or
    implementations.
  \item Formalized graphical abstract categorial grammars.
    \begin{itemize}
    \item Helps solve multiple constraints elegantly.
    \item Facilitates discussion and inquiry into compositions of ACGs.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Make a closer examination of the relationship between extrinsic
    and pangraphical languages in G-ACGs.
  \item Determine whether ACG object languages are closed on intersection
    or not.
  \item Build a wide-coverage G-ACG and study the patterns used in the
    graphical architecture.
  \item Try to express the multiple independent polarized features of IGs
    using G-ACGs.
  \item Assess the applicability of G-ACG architectures at other levels
    of description, e.g. discourse structure.
  \end{itemize}
\end{frame}
